---
title: "Analysis Report"
author: "Francisco"
date: "2024-07-23"
output: html_document
---
##Importing back .csv file for analysis
```{r}
cleandata <- read_csv("cleandata.csv")
```
##Creating Mode function in R
Mode <- function(x, na.rm = FALSE) {
  if(na.rm){
    x = x[!is.na(x)]
  }

  ux <- unique(x)
  return(ux[which.max(tabulate(match(x, ux)))])
}
##Performing summary analysis
```{r}
Mode(cleandata$day_of_week)
seconds_to_period(mean(cleandata$ride_length))
seconds_to_period(max(cleandata$ride_length))
member_status_summary <- 
    cleandata %>%
    group_by(member_casual) %>%
    filter(ride_length > 0) %>% 
    summarize(average_ride_length=seconds_to_period(round(mean(ride_length))),
              min_ride_length=seconds_to_period(round(min(ride_length), 2)),
              max_ride_length=seconds_to_period(round(max(ride_length))),
              most_common_day=Mode(day_of_week),
              most_common_start_station=Mode(start_station_name),
              most_common_end_station=Mode(end_station_name),
              ride_count = n()) %>%
    ungroup() %>%
    mutate(percent = label_percent()(ride_count/sum(ride_count)))
  day_of_week_summary <-
      cleandata %>% 
      group_by(day_of_week) %>% 
      filter(ride_length > 0) %>% 
      summarize(average_ride_length=seconds_to_period(round(mean(ride_length),1)),
                num_of_rides = n()) %>% 
      arrange(desc(num_of_rides))
```
##Visualizing
```{r echo=TRUE}
cleandata %>%
  mutate(weekday = wday(started_at, label = TRUE)) %>%
  group_by(member_casual, weekday) %>%
  summarise(number_of_trips = n()
    ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, weekday) %>%
  ggplot(aes(x = weekday, y = number_of_trips, fill = member_casual)) +
  geom_col(position = "dodge")

cleandata %>%
  mutate(weekday = wday(started_at, label = TRUE)) %>%
  group_by(member_casual, weekday) %>%
  summarise(number_of_rides = n(),
            average_duration = mean(ride_length, na.rm = TRUE)) %>%
  arrange(member_casual, weekday) %>%
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
```
##Exporting .csv files for further analysis
```{r}
write_csv(day_of_week_summary, "dayofweeksumm.csv")
write_csv(member_status_summary, "memberstatus.csv")
```